---
title: "Casino floor mix optimization"
author: "Andreose E, Caldara G, Comotto F, Scardovelli M"
output:
  html_document:
    df_print: paged
---

Import libraries and dataset
```{r}
library(tidyverse)
library(lubridate)
library(caret)
library(linprog)
options(scipen = 999)
slot <- readxl::read_excel("data/Lucky Duck Entertainment revenue 2013.xls", na = c(".", "NA", "NaN"))
str(slot)
```

Create dataset to be used for further analysis, i.e. aggregate slots by categories (MachineType+Denomination) for each month and section in each casino.
```{r}
df <- slot %>%
  group_by(Month, Casino, Section, MachineType, Denomination) %>%
  summarise(numero_macchine    = sum(NoMachines),
            ricavi_totali      = sum(GrossRevenue),
            ricavo_unitario    = ricavi_totali/numero_macchine,
            giocate_totali     = sum(Plays),
            giocate_unitarie   = round(giocate_totali/numero_macchine),
            ricavo_per_giocata = ricavi_totali/giocate_totali) %>%
  arrange(Month, Section, MachineType, Denomination) %>%
  mutate(tipo = paste0(MachineType, "_", Denomination)) 
head(df)

# df %>% ggplot(aes(x = giocate_totali, y = ricavo_per_giocata)) +
#   geom_point()
# df[df$"ricavo_per_giocata" > 5,]
```

Temporary: use only ...
```{r}
df <- filter(df, Month == ymd("2011-09-01")) #!= 
```

Some preprocessing
```{r}
numslots.section <- df %>%
  group_by(Month, Casino, Section) %>%
  summarise(sum = sum(numero_macchine),
            ricavo = sum(ricavi_totali)) %>%
  #il mutate perde un livello dopo il summarise
  mutate(sum_casino = sum(sum), prop_per_mese = round(sum / sum(sum), 3))


# ggplot(data = numslots.section, aes(x = sum, y = prop_per_mese)) +
#   geom_point(aes(color = Casino, size = ricavo)) +
#   ggtitle("Distribuzione della proporzione del numero di slot sul totale del casino per ogni mese") +
#   xlab("Numero slot per sezione (per mese)") +
#   ylab("Proporzione di slot in ogni mese") #+ facet_wrap(~Section)

#Ottieni numero di slot presenti in ogni casino per mese da usare come upper bound
numslots.casino <- NULL #a cosa serve ? 
nslots <- df %>%
  group_by(Month, Casino) %>%
  summarise(sum = sum(numero_macchine)) %>%
  arrange(Month) #ensure order

nslots <- nslots$sum #numero di slot per casino e mese
```


Definition of the objective function using profit coefficients
```{r}
f.obj <- round(df$ricavo_unitario)
```

Costruzione dei vincoli
Primo vincolo: numero massimo di slot per ogni casino e ogni mese. 
```{r}
#First create dummy var for Casino and month, used to activate each constraint depending on the casino and month
f.A1 <- t(predict(dummyVars(~ Casino, data = df), newdata = df))
#View(rbind(t(df[, "Casino"]), t(df$Month), f.A)) #Check results

#with one month don't require " : I(factor(Month)" after Casino


#Coefficients
b1 <- nslots #month1-> casino1, month1 ->casino2, month2 -> casino1, ...
f.b1 <-  b1 #vector of right-hand side values - upper bounds of total number of slots per month
f.dir1 <- rep("<=", length(f.b1)) # direction of the inequalities

sol1 <- solveLP(f.obj, f.b1, f.A1, maximum = TRUE, const.dir = f.dir1) #solver 
sol1bis <- lp(direction = "max", objective.in = f.obj, const.mat = f.A1, const.dir = f.dir1, const.rhs = f.b1)

sol1$con
sol1$solution
```

Secondo vincolo: la proporzione del tipo (MachineType+Denomination) di slot all'interno di ciascuna sezione dev'essere rispettato per ogni mese. Ovvero, non ci puo' essere un numero di slot di una categoria estremamente alto, ma deve rispettare dei vincoli 'di proporzione'.
Questo vincolo include il fatto che non ci possano essere sezioni vuote.
##Vincolo 2
#Soluzione Fede
Cambia proporzione minima e massima per ogni mese. 
```{r}
f.A2 <- t(predict(dummyVars(~ Section : Casino, data = df), newdata = df))
#with one month don't require " : factor(Month)" after Casino

# Duplicate every row, needed because for each section/month/casino we need a lower and upper bound. 
# Retain order --> ob1-low, obs1-up, obs2-low, ...)
f.A2 <- as.tibble(f.A2) %>% slice(rep(1:n(), each = 2)) %>% as.matrix()

#I vincoli sono attivati nel seguente modo 
# mese1-casino1-sez1-low,
# mese1-casino1-sez1-up,
# mese1-casino1-sez2;
# mese1-casino2;
# mese2;...
# quindi i primi 8 (4 sez * 2 up/low) vincoli sono associato ad un medesimo upper e lower bound, legato al primo valore di nslots.

#nslots contiene il numero di slot per ogni casino e mese, alternato per casino (mese1 casino1, mese1 casino2, mese2 casino1, ...)
# combine nslots in two columns, where rows are months and columns are casino (first for Aries)
(numslots <- matrix(nslots, ncol = 2, byrow = T))
f.b2 <- c()


month_prop <- numslots.section %>%
  group_by(Month, Casino) %>% summarise(prop_min = min(prop_per_mese),
                                        prop_max = max(prop_per_mese))

b2 <- cbind(nslots, month_prop$prop_min, month_prop$prop_max)
colnames(b2) <- NULL #altrimenti errore nel solver 
# to check correctness, compare with 
#View(numslots.section)
f.b2 <- c()
# probabilmente si puo' fare senza il ciclo for, ma come??
for (i in 1:nrow(b2)) {
  f.b2 <- c(f.b2, rep(round(c(b2[i,1]*b2[i,2], b2[i,1]*b2[i,3])), 4))
}

f.dir2 <- rep(c(">=", "<="), length(f.b2)/2) # direction of the inequalities

sol2 <- solveLP(f.obj, f.b2, f.A2, maximum = TRUE, const.dir = f.dir2) #solver 
sol2bis <- lp(direction = "max", objective.in = f.obj, const.mat = f.A2, const.dir = f.dir2, const.rhs = f.b2)

summary(sol2)
sol2$con
```

#Soluzione Gabri
Proporzione minima e massima per la serie storica intera diventano le soglie per tutti i mesi (cioe' fisse). 
```{r}
# Da notare, rispetto alla soluzione precedente, l'ordine in dummyVars differente
f.A2 <- t(predict(dummyVars(~ Casino : Section, data = df), newdata = df))
# see previous ": factor(Month)". In this case i think that putting Section or Casino first doesn't change the result.

# Duplicate every row, needed because for each section/month/casino we need a lower and upper bound. 
# Retain order --> ob1-low, obs1-up, obs2-low, ...)
f.A2 <- as.tibble(f.A2) %>% slice(rep(1:n(), each = 2)) %>% as.matrix()

#I vincoli sono attivati nel seguente modo 
# mese1-casino1-sez1-low,
# mese1-casino1-sez1-up,
# mese1-casino2-sez1;
# mese1-casino1-sez2;
# mese2;...

#nslots contiene il numero di slot per ogni casino e mese, alternato per casino (mese1 casino1, mese1 casino2, mese2 casino1, ...)
# combine nslots in two columns, where rows are months and columns are casino (first for Aries)
numslots <- matrix(nslots, ncol = 2, byrow = T)
f.b2 <- c()
# probabilmente si puo' fare senza il ciclo for, ma come??
for (i in 1:nrow(numslots)) {
  f.b2 <- c(f.b2, rep(round(c(numslots[i,1]*.2, numslots[i,1]*.32, numslots[i,2]*0.18, numslots[i,2]*0.35)), 4))
}
f.dir2 <- rep(c(">=", "<="), length(f.b2)/2) # direction of the inequalities

sol2 <- solveLP(f.obj, f.b2, f.A2, maximum = TRUE, const.dir = f.dir2) #solver 
sol2bis <- lp(direction = "max", objective.in = f.obj, const.mat = f.A2, const.dir = f.dir2, const.rhs = f.b2)

summary(sol2)
sol2$con

#To see solution on the original dataset
# df[which(sol2$solution != 0),] %>%
#   arrange(Casino, tipo) %>%
#   View()
```

Vincolo 3 - Preprocessing
```{r}
#Calcola i bin per le giocate totali del casino Libra, per differenziare dai bin di Aries
quartili.aries <- c(quantile(df[df$Casino == "Aries", ]$"giocate_totali", probs = seq(0, 1, by = 0.25)))
quartili.libra <- c(quantile(df[df$Casino == "Libra", ]$"giocate_totali", probs = seq(0, 1, by = 0.25)))
#Check dataset ordering --> arrange(Month, Section, MachineType, Denomination, Casino)
head(df, n = 100)
head(arrange(df, Month, Section, MachineType, Denomination, Casino), n = 100) #ok, e' cosi'
# split dataset based on casino, to differentiate bins
aries.idx <- which(df$Casino == "Aries")
aries <- df[aries.idx,]
libra <- df[-aries.idx,]
# Create bins based on total plays, include lowest value, and display digits instead of scientific notation
aries$bin_giocate <- cut(aries$giocate_totali, breaks = quartili.aries, include.lowest = T, dig.lab=10)
libra$bin_giocate <- cut(libra$giocate_totali, breaks = quartili.libra, include.lowest = T, dig.lab=10)
# merge the two casino into the original dataset, with the additional column bin_giocate
df <- rbind(aries, libra) %>% arrange(Month, Section, MachineType, Denomination, Casino)
# reorder columns, set bin_giocate next to giocate_totali
df[c(1:9, 13, 10:12)]
# Confirm ordering is the same as before
all(which(df$Casino == "Aries") == aries.idx)

bin_stats <- df %>% 
  group_by(bin_giocate, Casino) %>%
  summarise(minimo_macchine = min(numero_macchine),
            massimo_macchine = max(numero_macchine),
            media_macchine = mean(numero_macchine),
            mediana_macchine = median(numero_macchine),
            totale_macchine = sum(numero_macchine)) %>%
  #arrange(Casino, desc(bin_giocate)) #non funziona bene come ordinamento
  arrange(Casino, media_macchine) #meglio

#Non funzia
par(mfrow = c(2, 4))
for (b in unique(df$bin_giocate)) {
  ggplot(data = df[df$bin_giocate == b,], aes(x = "b", y = numero_macchine)) + 
    geom_boxplot()
}
par(mfrow = c(1,1))
```


Vincolo 3 - vincola il numero di slot di ogni categoria tra un min e max calcolato per gruppi di slot che condividono uno stesso range di numero di giocate.
```{r eval=FALSE, include=FALSE}
# create all combinations of ...
A3 <- t(predict(dummyVars(~ bin_giocate : factor(Denomination) : MachineType : Casino : Section , data = df), newdata = df))
#see previous ": factor(Month)"

# we must remove empty rows (that is no observation matches the combination, 
# think for example about bins belonging to another Casino, they will still be computed but with no matching observation)
# exclude all rows that have all zeros (sum is not more than zero)
A3 <- A3[rowSums(A3[,])>0, ]
# check dimension, it should have a number of rows equal to the number of observation in the dataset 
# since we are constraining each Month/Casino/Section/category(tipo) of slots, that is how our dataset is constructed
dim(A3) #bingo!
# Nowm duplicate every row, needed because we need to set a lower and upper bound for the number of slots. 
# Retain order --> ob1-low, obs1-up, obs2-low, ...)
f.A3 <- as.tibble(A3) %>% slice(rep(1:n(), each = 2)) %>% as.matrix()   
#l'errore sta nella creazione della matrice A

b <- c()
for (i in 1:nrow(df)) {
  #get the correct row index to identify the bin reference
  idx <- match(df[i,"bin_giocate"], bin_stats$bin_giocate)
  # create vector of constraint coefficient, lower then upper bounds
  b <- c(b, bin_stats[idx, "minimo_macchine"], bin_stats[idx, "massimo_macchine"])
}
#result b vector is list, because bin_stats is a dataframe (tibble to be precise)
f.b3 <- unlist(b, use.names=FALSE)

f.dir3 <- rep(c(">=", "<="), dim(df)[1]) # direction of the inequalities
# length should be equal to the length of the constraints' coefficient
length(f.dir3)

sol3 <- solveLP(f.obj, f.b3, f.A3, maximum = TRUE, const.dir = f.dir3) # Va in loop, quello sotto no. 
sol3bis <- lp(direction = "max", objective.in = f.obj, const.mat = f.A3, const.dir = f.dir3, const.rhs = f.b3)

#sol3$con
summary(sol3)
sol3bis
```

#PROVA FEDERICO

```{r}
f.A3 <- diag(nrow(df))
f.A3 <- as.tibble(f.A3) %>% slice(rep(1:n(), each = 2)) %>% as.matrix()

b <- c()
for (i in 1:nrow(df)) {
  #get the correct row index to identify the bin reference
  idx <- match(df[i,"bin_giocate"], bin_stats$bin_giocate)
  # create vector of constraint coefficient, lower then upper bounds
  b <- append(b, bin_stats[idx, "minimo_macchine"])
  b <- append(b, bin_stats[idx, "massimo_macchine"])
}

f.b3 <- unlist(b, use.names=FALSE)
f.dir3 <- rep(c(">=", "<="), dim(df)[1]) # direction of the inequalities

#check if A3 is equal to f.A3 -> all(A3 == f.A3) FALSE
```



Unisci vincoli - soluzione totale finora
```{r}
f.A <- rbind(f.A1, f.A2, f.A3)
f.b <- c(f.b1, f.b2, f.b3)
f.dir <- c(f.dir1, f.dir2, f.dir3)

sol <- solveLP(f.obj, f.b, f.A, maximum = TRUE, const.dir = f.dir) #solver 
sol.bis <- lp(direction = "max", objective.in = f.obj, const.mat = f.A, const.dir = f.dir, const.rhs = f.b)

sol$solution
sol.bis
summary(sol)
shadow_price <- sol$con
```

#Sensitivity Analysis
When solving an LP problem we assume that values of all model coefficients are known with certainty. Such certainty rarely exists. Sensitivity analysis helps answer questions about how sensitive the optimal solution is to changes in various coefficients in a model.  
We made Sensitivity Analysis only for September: Each consideration, with the right changes, could be extended to the other months.

First of all: Maximum Number of machines in each Casino
```{r}
vincolo1 <- shadow_price[1:2,]
knitr::kable(vincolo1) 
```
Observations:  
- Each constraint is binding: the total number of machines for each Casin? is equal to the previous configuration;  
- For each extra machines in Aries (first constraint) the values of Obj. Function increases by 11095???. The maximun number of extra machines is 38, so, without considerating the purchase cost and assuming all others coeff. remain constant, the Obj. Function could increase by $$11095??? * 38 = 421610??? $$.  
- For each extra machines in Libra (Second constraint) the values of Obj. Function increases by 11562???. The maximun number of extra machines is 18 so, without considerating the purchase cost and assuming all others coeff. remain constant, the Obj. Function could increase by $$11562??? * 18 = 208116??? $$.  

Conclusion: If we have to increase the number of machines for only one Casin?, we'll choose Aries

Second one: Number of machines in each section (proportion) 
We remind that each constraint has lower and upper bound.
```{r}
vincolo2 <- shadow_price[3:18,]
knitr::kable(vincolo2) 
```
Let's look the upper bound to know which section have the maximum number of machines.  
Only constraints n. 14 n. 16 are binding, this means:  
- In Aries the Section that receives more machines (0.18 of the total) is "Restaurant Plaza" (constraint 7);  
- In Libra the Section that recerives more machines (0.35 of the total) is "Interior" (constraint 6).  

Let's look the lower bound to know which section have the minimum number of machines.  
In this case constraints n. 5,7,9 received the minimun number of machines: the number of machines is just sufficient to cover the lower bound, this means:  
- In Aries one Section, Entrace, receives less machines that the other ones;  
- In Libra two Sections, Boundary and Entrance, receive less machines that the other ones.  

Also in this case we can see the dual (shadow price) and dual.reg to decide which section could have more machines.

Last one: Number of machines of each category in each section
We remind that each constraint has lower and upper bound.
```{r}
vincolo3 <- shadow_price[19:length(shadow_price[,1]),]
rownames(vincolo3) <- 1:length(vincolo3[,1])
knitr::kable(head(vincolo3))
```

Let's look the upper bound to know which category in each section have the maximum number of machines.
```{r}
knitr::kable(vincolo3[vincolo3$actual == vincolo3$bvec & vincolo3$dir == "<=", ])
```
Ok queste sono le categorie che in ogni sezione hanno ricevuto il numero massimo di macchine: riesci a farmi una funzioncina che mi riconduca alle macchine del dataframe di partenza? Forse ti conviene usare il dataframe che ho costruito prima "vincolo 3" che contiene solamente le informazioni relative al terzo vincolo (upper e lower bound). Ho riordinato le righe da 1 a 156 con rownames pensando servisse. 

...Second Part...
```{r}
knitr::kable(sol$allvar)
```
